---
title: "Simple document"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

## Import some data

I want to import the file 'FSA_litters.csv'.

```{r}
litters_df = read_csv("data/FAS_litters.csv")
```

Yay! I imported the dataset. Now I want better variables names, preferably in snake case.

```{r}
names(litters_df)

litters_df = janitor::clean_names(litters_df)
```

Yay! Now I have better names. Let's look at the dataset. 

```{r}
litters_df

head(litters_df)
tail(litters_df)
```

One Jeff uses a lot to view litters_df in new window you can scroll through without any output:
```{r, eval = FALSE}
view(litters_df)
```

Here's 'skimr':

```{r}
skimr::skim(litters_df)
```

## Arguments in 'read_csv'

```{r}
litters_df =
    read_csv(
      "data/FAS_litters.csv",
      skip = 5,
      col_names = FALSE,
      na = "Low8"
    )
```

## Parsing columns

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(
    `GD of Birth` = col_integer()
  )
)
```

## Reading from Excel

```{r}
mlb11_df = read_excel("data/mlb11.xlsx")
```

LotR Words is next.

```{r}
fellow_df = read_excel("data/LotR_Words.xlsx", range = "B3:D6")
```


## Read a SAS file

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
```


## Why to never use 'read.csv'

It will always print things as a dataframe instead of a tibble. Read_csv is preferred. 'read.csv' doesn't give 'tibble' and that's really bad. 

```{r}
litters_df_base = read.csv("data/FAS_litters.csv")
```


## Exporting data 

There's a good way to write data out

```{r}
write_csv(fellow_df, "data/fellowship_words.csv")
```

