---
title: "Tidy Data"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

## `pivot_longer`

Load the PULSE data.
```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Wide format to long format ... 

```{r}
pulse_data_tidy = 
  pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)
  )
```

Rewrite, combine, and extend (to add a mutate)
```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  relocate(id, visit) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))
```

(don't use `gather()`, just use `pivot_longer()`)

## `pivot_wider`

Make up some data!

```{r}
analysis_df = 
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 8, 3.5, 4)
  )

analysis_df %>% 
  pivot_wider(
    names_from = "time",
    values_from = "mean"
  ) %>% 
knitr::kable()
```

(don't use `spread()`, just use `pivot_wider()`)

## Binding rows

Using the LotR data.

First step: import each table. 

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")

```

Bind all the rows together (never use `rbind()` always use `bind_rows()`)

```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words"
  ) %>% 
  mutate(race = str_to_lower(race)) %>% 
  relocate(movie) 
```

## Joining datasets

Import and clean the FAS datasets. This imports and cleans litters and pups data.

```{r}
pups_df =
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))
  
litters_df =
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  relocate(litter_number) %>% 
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>% 
  mutate(dose = str_to_lower(dose))
```

Next up, time to join them!

```{r}
fas_df = 
  left_join(pups_df, litters_df, by = "litter_number") %>% 
  arrange(litter_number) %>% 
  relocate(litter_number, dose, day_of_tx)
```

